---
title: "Take-home_Ex03"
author: "Mohamad Hafez Sidek"
date: "June 9, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
  freeze: true
  message: false
---

**Installing R Packages**

```{r}

pacman::p_load(ggraph,dendextend,tidygraph,jsonlite,dplyr,lubridate,igraph,ggplot2,gridExtra)
```
---

Load JSON data

---
```{r}

json_data <- fromJSON('data/MC3/mc3.json')
```
---

Extract the nodes and edges

---
```{r}

nodes <- json_data$nodes
edges <- json_data$links
```
---

Data Processing

---
```{r}

# Ensure 'id' columns and edge 'source'/'target' columns are character type
nodes$id <- as.character(nodes$id)
edges$source <- as.character(edges$source)
edges$target <- as.character(edges$target)

# Convert date fields in nodes
nodes$`founding_date` <- as.Date(nodes$`founding_date`, format = "%Y-%m-%d")
nodes$`_last_edited_date` <- as.Date(nodes$`_last_edited_date`, format = "%Y-%m-%d")
nodes$`_date_added` <- as.Date(nodes$`_date_added`, format = "%Y-%m-%d")

# Convert date fields in edges
edges$`start_date` <- as.Date(edges$`start_date`, format = "%Y-%m-%d")
edges$`end_date` <- as.Date(edges$`end_date`, format = "%Y-%m-%d")

```
---


---
```{r}


```
***Problem Statement 1:***
Identify the network associated with SouthSeafood Express Corp and visualize how this network and competing businesses change as a result of their illegal fishing behavior.

```{r}


```
***Analysis***


----

***Problem Statement 2:***
 Which companies benefited from SouthSeafood Express Corp legal troubles? 
```{r}


```
***Analysis***


----

***Problem Statement 3:***
Are there any other suspicious transactions that may be related to illegal fishing?

---

Data Processing

---
```{r}

# Convert nodes to a data frame for filtering purposes
nodes_df <- as.data.frame(nodes)

# Filter the dataset for companies involved in fishing, sea food, and seafood
fishing_companies <- nodes_df %>%
  filter(grepl("fish|sea food|seafood", TradeDescription, ignore.case = TRUE))

# Remove rows with missing or non-finite revenue values
fishing_companies <- fishing_companies %>%
  filter(!is.na(revenue) & is.finite(revenue))

# Identify top outliers by revenue (95th percentile)
top_outliers <- fishing_companies %>%
  filter(revenue > quantile(revenue, 0.95, na.rm = TRUE))

# Extract the IDs of the top outliers
top_outlier_ids <- top_outliers$id

# Convert edges to a data frame for filtering purposes
edges_df <- as.data.frame(edges)

# Filter edges to include only those related to top outliers
related_edges <- edges_df %>%
  filter(source %in% top_outlier_ids | target %in% top_outlier_ids)

# Create an edge list for igraph
edge_list <- as.matrix(related_edges[, c("source", "target")])

```

----

Create graph

----

```{r}

# Create graph object
g <- graph_from_edgelist(edge_list, directed = FALSE)

# Add vertex attributes for revenue and country
V(g)$revenue <- ifelse(V(g)$name %in% top_outliers$id, top_outliers$revenue[match(V(g)$name, top_outliers$id)], NA)
V(g)$country <- ifelse(V(g)$name %in% top_outliers$id, top_outliers$country[match(V(g)$name, top_outliers$id)], NA)

# Set node sizes and colors
V(g)$size <- ifelse(V(g)$name %in% top_outliers$id, 10, 5)  # Increase size for top outliers
V(g)$color <- ifelse(V(g)$name %in% top_outliers$id, "skyblue", "orange")  # Different colors for top outliers
V(g)$shape <- ifelse(V(g)$name %in% top_outliers$id, "square", "circle")  # Different shapes for top outliers
E(g)$color <- adjustcolor("gray", alpha.f = 0.5)  # Transparent edges



# Plot the network graph
plot(g, vertex.size = V(g)$size, vertex.label.cex = ifelse(V(g)$name %in% top_outliers$id, 0.8, 0.6), 
     vertex.color = V(g)$color, vertex.shape = V(g)$shape,
     vertex.label = NA, 
     edge.color = E(g)$color, layout = layout_with_fr, 
     main = "Network Graph of Top Revenue Outliers")

```
***Analysis***
Potential Implications for Suspicious Transactions Related to Illegal Fishing:
Role of Central Nodes:

***Point 1**
The central nodes (skyblue squares) with many connections could be orchestrating significant activities within the network.
Their prominent role and high connectivity suggest they might be key players in suspicious transactions related to illegal fishing.


***Point 2**
Dense Clusters:The dense clusters around these central nodes indicate high levels of transactional activity.
This could be a sign of coordinated efforts to obscure illegal fishing activities through multiple layers of transactions.
Isolation of Transactions:

***Point 3**
The presence of smaller, isolated subgroups suggests attempts to hide certain transactions from the main network.
These isolated groups might be used to carry out suspicious activities without drawing attention.


***Point 4**
Systematic Patterns:The repeated patterns of smaller clusters connected to central nodes indicate a systematic approach to transactions.
This systematic behavior could be indicative of coordinated efforts to manage and conceal illegal activities related to fishing.

***Conclusion:***
Based on the analysis of the network graph, there are several visual hints that suggest the possibility of suspicious transactions related to illegal fishing. The prominent role of the central nodes, the dense clusters of transactions, the isolated subgroups, and the systematic patterns all point towards potential coordinated efforts to obscure illegal activities. Further investigation into the nature of these transactions and the entities involved is warranted to confirm these suspicions.


