---
title: "Take-home_Ex01"
author: "Mohamad Hafez Sidek"
date: "April 26, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
  freeze: true
---

**Installing R Packages**

```{r message=FALSE, warning=FALSE}

pacman::p_load(tidyverse,ggplot2,readr,dplyr,lubridate,stringr)
```

**Load and Consolidate Data**

```{r message=FALSE, warning=FALSE}

realis_1 <- read_csv("data/ResidentialTransaction20240308160536.csv")
realis_2 <- read_csv("data/ResidentialTransaction20240308160736.csv")
realis_3 <- read_csv("data/ResidentialTransaction20240308161009.csv")
realis_4 <- read_csv("data/ResidentialTransaction20240308161109.csv")
realis_5 <- read_csv("data/ResidentialTransaction20240414220633.csv")
realis <- bind_rows(realis_1, realis_2, realis_3, realis_4, realis_5)
```

**Amend Character Types**

```{r message=FALSE, warning=FALSE}

realis$`Sale Date` <- as.Date(realis$`Sale Date`, format = "%d %b %Y")
realis$`Nett Price($)` <- as.numeric(gsub("[^0-9.]", "", realis$`Nett Price($)`))

```

**Filter first quarter of 2024 only**

```{r message=FALSE, warning=FALSE}

realis2024Q1 <- realis %>%
  filter(year(`Sale Date`) == 2024 & quarter(`Sale Date`) == 1)

```

**Group the data by Property Type and calculate the sum Number of Units for each group**

```{r message=FALSE, warning=FALSE}

Number_of_Units_by_Property_Type <- realis2024Q1 %>%
  group_by(`Property Type`) %>%
  summarise(`Total Number of Units` = sum(`Number of Units`))
```

**Reorder levels of Property Type based on Total Number of Units**

```{r message=FALSE, warning=FALSE}

Number_of_Units_by_Property_Type <- Number_of_Units_by_Property_Type %>%
  arrange(desc(`Total Number of Units`)) %>%
  mutate(`Property Type` = factor(`Property Type`, levels = `Property Type`))
```

**Calculate cumulative percentage of total units by Property Type**

```{r message=FALSE, warning=FALSE}
Number_of_Units_by_Property_Type <- Number_of_Units_by_Property_Type %>%
  mutate(Cumulative_Percentage = cumsum(`Total Number of Units`) / sum(`Total Number of Units`) * 100)
```

**Plot Pareto chart**

```{r message=FALSE, warning=FALSE}
ggplot(data = Number_of_Units_by_Property_Type, aes(x = `Property Type`)) +
  geom_bar(aes(y = `Total Number of Units`), stat = "identity", fill = "grey70") +
  geom_line(aes(y = Cumulative_Percentage * max(`Total Number of Units`) / 100), color = "red", size = 1) +
  geom_point(aes(y = Cumulative_Percentage * max(`Total Number of Units`) / 100), color = "blue", size = 2) +
  scale_y_continuous(sec.axis = sec_axis(~./max(.) * 100, name = "Cumulative Percentage")) +
  theme_minimal() +
  labs(y = "Total Number of Units", x = "Property Type", title = "Pareto Chart of Total Number of Units by Property Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```

**\*\* EDA 1 \*\***

Pareto chart distinctly illustrates the imbalance in property types, with Condominiums and Apartments not only dominating the market but also highlighting the Pareto principle---these categories alone account for a disproportionate 75% of the total units.

This suggests higher demand or greater supply in these sectors. As the chart progresses towards more exclusive properties like Terrace Houses, Semi-Detached, and Detached Houses, there's a notable decline in unit numbers, hinting at lesser availability or possibly higher pricing brackets for such properties.

This trend is significant for stakeholders, indicating a potential focus area for both market demand and real estate development strategies.

**Create new columns**

```{r message=FALSE, warning=FALSE}
#Determine the address unit level.
realis2024Q1 <- realis2024Q1 %>%
  mutate("Floor Level" = if_else(str_detect(Address, '#'),
                                   str_trim(str_extract(str_split(Address, "#", simplify = TRUE)[, 2], "^[^\\-]*")),
                                   "N.A."))

#Determine the address unit level category (Low, middle or high).
realis2024Q1 <- realis2024Q1 %>%
  mutate(`Floor Level Category` = case_when(
    `Property Type` %in% c("Terrace House", "Semi-Detached House", "Detached House") ~ "N.A.",
    `Floor Level` == "N.A." ~ "Low",
    as.integer(`Floor Level`) <= 5 ~ "Low",
    as.integer(`Floor Level`) <= 10 ~ "Middle",
    TRUE ~ "High"
  ))


```

**Plot Boxplot**

```{r message=FALSE, warning=FALSE}
# Reorder levels of 'Property Type' alphabetically
realis2024Q1$`Property Type` <- factor(realis2024Q1$`Property Type`, levels = c("Apartment","Condominium","Terrace House", "Semi-Detached House","Detached House","Executive Condominium"))

# Reorder levels of 'Floor Level Category' by "High", "Middle", "Low"
realis2024Q1$`Floor Level Category` <- factor(realis2024Q1$`Floor Level Category`, levels = c("High", "Middle", "Low"))

# Create the boxplot using ggplot2
ggplot(realis2024Q1, aes(x = `Property Type`, y = `Unit Price ($ PSM)`, fill = `Floor Level Category`)) +
  geom_boxplot() +
  scale_fill_manual(values = c("N.A." = "gray", "Low" = "blue1", "Middle" ="blue3", "High" = "blue4")) +
  labs(x = "Property Type", y = "Unit Price ($ PSM)", fill = "Floor Level Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

**\*\* EDA 2 \*\* Insert Analysis Here**

The boxplot analysis highlights several significant findings, with a focus on median prices:

1.  Terraced, semi-detached, and detached houses exhibit lower median prices in contrast to vertical living spaces such as apartments ,and condominiums. This discrepancy suggests a potential premium associated with apartment living, possibly attributed to location advantages, particularly in central urban areas.

2.  Condominiums generally boast higher median prices compared to apartments, indicative of additional amenities or sought after locations that elevate their market value.

3.  Executive condominiums stand out with the lowest median price among all property types, implying a level of affordability. This could be due to the regulatory price controls.

4.  Further exploration into median prices based on property type and floor levels reveals that higher floors command a premium for condominiums and apartments.

5.  In contrast, executive condominiums exhibit a more uniform pricing pattern across different floors, in comparison to apartments and condominiums.

**\*\* Summary\*\***

The analysis of property types using Pareto charts sheds light on the distribution within the market. Condominiums and apartments emerge as prominent, aligning with the Pareto principle by occupying a significant portion of total units. This suggests either heightened demand or extensive supply in these segments.

This finding is further corroborated by the examination of property types using boxplots. It highlights the preference for apartment living over traditional housing types, emphasizing the premium associated with such accommodations.

In essence, these analyses provide valuable insights into the property market, emphasizing the dominance of condominiums and apartments and their impact on demand and pricing dynamics.
